Numerical Investigation of Viscous Flow Instabilities in Multiphase Heterogeneous Porous Media

### 2018-02-08 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
updates

%%%%%%%%%%%%%
\section{Conclusions}\label{Section:Conclusion}
% Short (1 paragraph) summary of the paper with focus on the main theoretical aspects that were proved by our work in the present manuscript;
The purpose of this paper is to numerically investigate the dynamics of viscous flow instabilities in porous media. Simulations involving formation and growth of viscous fingers were conducted with the CVFEM-based multi-fluid flow simulator Fluidity/IC-FERST model. The model is based upon a dual consistent pressure-velocity representation in CV and FEM spaces and a novel family of FE-pairs, \PN[n]{m} and \PNDG[n]{m}. Saturation and other scalar fields are discretised in CV space and may be fully discontinuous across finite element boundaries. %Material absolute permeability is piecewise constant in FE space to ensure the correct representation of the surface.
In order to strongly enforce discretised Darcy equations at the boundaries between elements and control volumes, a directional-weighted flux-limited was introduced to take into account discontinuities risen from both, numerical formulation (based on discontinuous-Galerkin finite element method, DGFEM) and control volumes within finite elements.

During immiscible multi-fluid flow displacements, the expected uniform front is perturbed if fluids have different densitites and/or viscositiesleading to an uneven front with elongations at the outside edge of the fluid interface

>>>>

%%%%%%%%%%%%%
\section{Conclusions}\label{Section:Conclusion}
% Short (1 paragraph) summary of the paper with focus on the main theoretical aspects that were proved by our work in the present manuscript;
During immiscible multi-fluid flow displacements, the expected uniform front becomes unstable due to: (a) density ratio; (b) viscosity ratio, and; (c) heterogeneity. If viscosity ratio is relatively large (VR$\ge$3), surface tension becomes weak and the interface between two fluids is unstable in the presence of tangential velocity discontinuities, leading to the formation of fingers \citep{saffman_1986}. Heterogeneity in the porous media domain  also leads to solution instabilities in the hyperbolic Buckley-Leverett problem \citep{langtangen_1992}. This paper aims to numerically investigate the dynamics of viscous flow instabilities in porous media triggered by permeability heterogeneity.

Simulations involving formation and growth of viscous fingers were conducted with the CVFEM-based multi-fluid flow simulator Fluidity/IC-FERST model. The model is based upon a dual consistent pressure-velocity representation in CV and FEM spaces and a novel family of FE-pairs, \PN[n]{m} and \PNDG[n]{m}. Saturation and other scalar fields are discretised in CV space and may be fully discontinuous across finite element boundaries. In order to strongly enforce discretised Darcy equations at the boundaries between elements and control volumes, a directional-weighted flux-limited was introduced to take into account discontinuities risen from both, numerical formulation (based on discontinuous-Galerkin finite element method, DGFEM) and control volumes within finite elements.

 is perturbed if fluids have different densitites and/or viscositiesleading to an uneven front with elongations at the outside edge of the fluid interface
